---
title: "Improved \\(K\\)-means algorithm"
author: "Houda Aiboud Benchekroun and Thomas Roiseux"
date: "`r Sys.Date()`"
output: pdf_document
---
\rule{\linewidth}{0.1pt}
\part{Improved algorithm}
\section{Regular algorithm}
Before writing the improved \(K\)-means, we just want to make a quick reminder on the regular one:
```{r kmeans, code = readLines("kmeans.R")}
```
On this base, we are now going to only rewrite the initialization of this algorithm, as the other parts remain the same in the improved one:
```{r kmeanspp, code = readLines("modified-kmeans.R")}
```
\section{Generation of test samples}
We are now going to use these data set to test our algorithms.
```{r samples, code = readLines("simulation.R")}
```

\clearpage
\part{Application to iris data}
\section{Preview of the data}
First, let's see what is this dataset.
```{r loadiris}
names(iris)
summary(iris)
head(iris)
plot(iris)
```
We can notice some linear correlation between the length and the width of the petals.
\section{Algorithm comparizon}
```{r algos_comp}
library(mclust)
Mclust(iris)
```
\section{Primary component analysis}
```{r pca}
library(FactoMineR)
library(factoextra)
#PCA(iris, scale.unit = TRUE, ncp = 5, graph = TRUE)
```