---
title: "Improved \\(K\\)-means algorithm"
author: "Houda Aiboud Benchekroun and Thomas Roiseux"
date: "`r Sys.Date()`"
output: pdf_document
---
\rule{\linewidth}{0.1pt}
\part{Improved algorithm}
\section{Regular algorithm}
Before writing the improved \(K\)-means, we just want to make a quick reminder on the regular one:
```{r kmeans, code = readLines("kmeans.R")}
```
On this base, we are now going to only rewrite the initialization of this algorithm, as the other parts remain the same in the improved one:
```{r kmeanspp, code = readLines("modified-kmeans.R")}
```
\section{Generation of test samples}
We are now going to use these data set to test our algorithms.
```{r samples, code = readLines("simulation.R")}
```

\clearpage
\part{Application to iris data}
\section{Preview of the data}
First, let's see what is this dataset.
```{r loadiris}
names(iris)
summary(iris)
head(iris)
tail(iris)
plot(iris)
```
We can notice some linear correlation between the length and the width of the petals.
Also, regarding the variable names, we can guess that the \texttt{Species} variable will be qualitative.
```{r corrmat}
library(corrplot)
cormat <- cor(iris[, -5])#Removing species for correlation
corrplot(cormat)
```
The correlation matrix shows that we have strong correlation
between the width of sepals and petals, and the same for the petals.
\section{Algorithm comparizon}
```{r algos_comp}
library(mclust)
Mclust(iris)
```
\section{Primary component analysis}
Primary component analysis requires to have only quantitative variables in our data set.
However, as the \texttt{iris} data set, has a column \texttt{Species} that is qualitative (it describes gives the name of the considered iris species), we will first drop this column and then perform on only the remaining variables.
```{r pca}
library(FactoMineR)
library(factoextra)
df <- iris[, !(names(iris) %in% c("Species"))]
pca <- PCA(df, scale.unit = TRUE, ncp = 5, graph = FALSE)
fviz_pca_ind(pca, geom.ind = "point",col.ind = iris$Species,
             palette = c("#00AFBB", "#E7B800", "#FC4E07"),
             addEllipses = TRUE,
             legend.title = "Groups"
)
```

We now have the individual graphs. We therefore notice that, even is the \texttt{Species} is qualitative,
all observations are grouped by species.

```{r corcercle}
fviz_pca_var(pca, col.var = "contrib", repel = TRUE,
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"))
variables <- get_pca_var(pca)
head(variables$contrib)
corrplot(variables$cos2, is.corr = FALSE)
```

With the correlation circles, we know how the original variables are correlated with the new ones:
these new variables are made with linear combination of the original ones.
Therefore, using the previous circle and the contribution table, we have two principal components:
\begin{enumerate}
\item The first primary component, here named \texttt{Dim1} holds mainly all the data of the petals, which can be interpreted as the petals surface
\item The second primary component, here named \texttt{Dim2} is mostly made of the width and the length of the sepals. This can be the sepals surface.
\end{enumerate}